Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies

### 2024-06-17 paper.tex
Update paper.tex
data is garbage so deleted
rovib and interference trivial so deleted

added more on sfg to explain why i want to compare (it is to motivate doing the spectroscopy)

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Singly Vibrationally Enhanced (SIVE) spectroscopy, a type of nonlinear hyper-Raman spectroscopy, is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
%alternate name is hyper-(S/D)FG... which is more interesting and descriptive than SIVE imo. lmk your thoughts
We derive selection rules and coherent interference effects in singly resonant SIVE (SR-SIVE).
The theory of SR-SIVE is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant SIVE (DR-SIVE), akin to a resonant hyper-Raman process. 
DR-SIVE is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.
The linenarrowing abilities of DR-SIVE are discussed.
We showcase the difference in selection rules of SR-SIVE and DR-SIVE through a study of cyanocobalamin, highlighting the sensitivity of these measurements on non-resonant and resonant hyper-Raman hyperpolarizabilities, respectively.
\end{abstract}

@@

We investigate properties of doubly resonant SIVE (DR-SIVE) to understand how coupling between vibrational and vibronic states can be measured through DR-SIVE.
DR-SIVE is shown to demand one and two-photon absorptions between the ground and excited state.
DR-SIVE is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}
Based upon these selection rules, we investigate SR-SIVE and DR-SIVE spectroscopy of cyanocobalamin, highlighting the applicability of these methods to non-model systems.

\section{Experimental Methods}
The ultrafast laser system has been described in detail elsewhere. \cite{RN278, Kaufman2024}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator  (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG ``DF2'', Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 were computer controlled through communication with 50 mm DC actuators (PMC). %todo: who tf makes PMC motors?
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X. %todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused into the sample in a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale. %@dan if we should change to amp lmk

The cyanocobalamin (CNCbl) [DOT Scientific inc., CAS no. 68-19-9] samples were prepared as drop casted thin films.
CNCbl was dissolved in methanol (MeOH) [Fisher Scientific, HPLC Grade, CAS no. 67-56-1] until a concentrated solution was formed (concentration $\sim$ 10 mg CNCbl/ mL MeOH). 
All chemicals were used as obtained without further purification.
The solution was deposited on a microscope coverslip and allowed to dry in a dessicator until a thin layer was formed.

\section{Properties of SIVE Spectroscopy}
Singly resonant SIVE (SR-SIVE) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between SR-SIVE spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we make the connections between SIVE and hyper-Raman scattering explicit and remark on the properties of SIVE that become apparent after inspecting the selection rules.

@@

Note that we choose the convention where an arbitrary electric field, assuming a plane wave, is written in Fourier components as 
\begin{equation}
	\vec{E}(\omega) = \frac{1}{2} \vec{E}_0 \exp(i(\vec{k}\cdot\vec{x} - \omega t)) + c.c.
\end{equation}
where $E_0$ is the pulse amplitude and $\vec{k}$ is a wavevector.
%maybe talk about maxwell wave eqn for e fields ?
Expressions for $\gamma_{ijkl}$ can be found by solving Liouville-von Neumann equation for an output density matrix element $\rho_{ij}$.
Generally, under the electric dipole approximation, the density matrix, $\bar{\rho}(t)$, is propagated through the Liouville-von Neumann equation as \cite{RN455}
\begin{equation}\label{matrix}
	\dot{\bar{\rho}}(t) = \frac{1}{i \hbar}[H_0 - \vec{\mu}\cdot \sum_j E_j(\vec{x},t), \bar{\rho}(t)] + \bar{\Gamma} \bar{\rho}(t)
\end{equation} %check if bar or double bar
where $H_0$ is the system (time independent) Hamiltonian, $\vec{\mu}$ is the electric transition dipole moment, $\vec{E}(\vec{x},t)$ is a time dependent electric field, and the elements of $\bar{\Gamma}$, $\Gamma_{ij}$, are dephasing rates. 
In the rotating frame, $\tilde{\rho}_{ij} = \rho_{ij} e^{-i \omega_{ij}t}$.
By solving \autoref{matrix} under steady state ($\dot{\bar{\rho}}(t) = 0$) in the rotating frame, expressions for $\gamma_{ijkl}$ can be readily obtained by multiplying the output dipole moment, $\mu_i$ by the final density matrix element $\tilde{\rho}_{ij}$.

\subsection{Vibrational SR-SIVE Selection Rules}
To make the connection between SR-SIVE and hyper-Raman scattering, we investigate the gross selection rules of SR-SIVE-IR.
By propagating density matrix elements in the steady state limit, the SIVE hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state density.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a SIVE experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1934, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
As such, any mode that is SIVE active must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is SIVE active.
This makes SIVE an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
Unlike SFG spectroscopy, which is dependent upon both the infrared and Raman activity of a mode (through symmetry constraints, this demands that the microscopic SFG output from centrosymmetric species vanishes),\cite{Shen90, Cotton} SIVE is a nonlinear, optically upconverted spectroscopy for any infrared active mode.

@@

To motivate the application of SIVE spectroscopy, we perform an approximate calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where $\alpha_{ij}$ is the Raman polarizability, N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

\subsection{Rovibrational SR-SIVE Selection Rules} %this can probably be deleted but wanted to add just in case...
Recent developments in nonlinear spectroscopy have demonstrated the feasibility of multidimensional infrared spectroscopy of gaseous and quasi-free species. \cite{Ziegler2018, Gronborg2022, RN325, Chen2024}
DOVE and 2D-IR have been used to isolate coupling between fundamentals, overtones and combination bands of gaseous species in several molecular systems. 
Since resonance hyper-Raman scattering spectroscopy was developed as a gas phase spectroscopy,\cite{RN515} we briefly investigate the properties of rovibrational SR-SIVE.
In the discussion that follows, to avoid integration over velocity dependent distributions, we assume the experiment is performed at 0 K.
The velocity dependence of output density matrix elements does not manipulate quantum mechanical selection rules. \cite{Bracamonte2003}

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the total angular momentum and $M$ is its projection onto the space fixed quantization axis).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, the matrix elements in \autoref{sivebeta} can be evaluated through the Wigner-Eckart theorem.
It is important to note that $\beta_{ijk}$ does not have complete index symmetry.\cite{Andrews1978}
Specifically, it has shown that $T_{q_1}^{(1)}$ fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)}$ and  $T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Altmann1982, Andrews1990}
%there is some fancy way to note that theres an index constraint but i dont remember the language... i think is fine
As such, its decomposition into irreducible tensor form is allowed only if the values of $q_2, q_3$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

\subsection{Quantitative SR-SIVE}
With the selection rules of SR-SIVE understood for vibrational and rovibrational spectroscopy, it becomes possible to extract quantitative information from SIVE spectra to compare to incoherent methods and other CMDS methods.
Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
In the method introduced by Levenson and Bloembergen (Bloembergen Interferometry Experiment), an internal standard interferes with the resonant lineshape.
The internal standard is often chosen to be C$_6$H$_6$ or C$_6$D$_6$, due to the strong Raman activity and measured $\chi^{(3)}$ of its ring breathing mode ($\nu_2$ in Herzberg notation). \cite{Levenson1974_2, RN351, RN345}
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Resonant lineshapes are also complicated by amplitude level interference between the sample and its substrate and/or sample cell windows, which must be accounted for to obtain quantitatively correct $\chi^{(3)}$ values. \cite{RN362, RN418}

@@

\subsection{Coherent Interference in SR-SIVE}
It is well known that nonlinear wave mixing processes which share the same time ordering and output frequency will interfere. \cite{RN135, Bonn2024}
Understanding interfering pathways is useful because they can, in some cases, eliminate output. \cite{RN342, RN135}
The WMEL diagrams presented in \autoref{fig:sivewmel} outline resonant SIVE processes when the infrared pulse $(\omega_1)$ is resonant with a vibrational mode. 
However, if $\omega_2$ becomes vibrationally resonant and $\omega_1$ is non-resonant, then other SIVE pathways appear.\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:sivewmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} may significantly interfere with DOVE-IR and DOVE-Raman spectra.
The time ordering in DOVE-IR and DOVE-Raman spectra is used to interpret anharmonicities, as DOVE-IR and DOVE-Raman destructively interfere when outputs are at the same frequency. \cite{RN135, RN324}
It is therefore important to understand how the corresponding SIVE pathways interfere.
To this end, we investigate how the SIVE-IR-II and SIVE-Raman pathways (\autoref{fig:sivewmel2}a, b, respectively) interfere under the rotating wave approximation.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{wmel2.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of interfering SR-SIVE pathways in the  $\vec{k}_4 = -\vec{k}_1 + \vec{k}_2 + \vec{k}_3$ phasematching geometry. 
	(a) is similar to the process diagrammed in \autoref{fig:sivewmel}, whereas (b) is similar to a coherent anti-Stokes Raman (CARS)-type process.}
	\label{fig:sivewmel2}
\end{figure}
\begin{widetext}
We employ the response function formalism of Mukamel for this analysis. \cite{RN287}
The third order response functions, $R^{(3)}$, for SR-SIVE-IR-II ($R^{(3)}_{S_{II}}$) and SR-SIVE-Raman ($R^{(3)}_{R}$) (\autoref{fig:sivewmel2}a, b), suppressing tensor indices, are 
\begin{subequations}
	\begin{equation} \label{mixing:a}
		R^{(3)}_{S_{II}} (t_2) = -\frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2)
	\end{equation}
	\begin{equation}\label{mixing:b}
		\begin{split}
			R^{(3)}_{R} (t_2) & = \frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2) = - R^{(3)}_{S_{II}} (t_2)\\
		\end{split}
	\end{equation}
\end{subequations}
By assuming that the fields which stimulate hyper-Raman scattering, $\vec{E}_1, \vec{E}_3$, interact simultaneously at some time $\tau$ after $\vec{E}_2$ interacts with the sample, the total output polarization, $\vec{P}^{(3)}(t)$, is written as \cite{Cho2001}
	\begin{equation}\label{polarization_SRSIVE_a}
		\begin{split}
			\vec{P}^{(3)}(t) &= \vec{P}_{S_{II}}(t) + \vec{P}_{R}(t)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) + R^{(3)}_{R} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) - R^{(3)}_{S_{II}} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau) \\
			&= 0
		\end{split}
	\end{equation}
It is clear the total output polarization for SR-SIVE-IR-II and SR-SIVE-Raman vanishes, or, SR-SIVE-IR-II and SR-SIVE-Raman perfectly destructively interfere.
\end{widetext}
The absence of SR-SIVE-IR-II pathways in ultrafast DOVE work is rationalized by this calculation. \cite{RN367, McDonnell2024}
In the case of the time ordered SR-SIVE-IR-I pathway, there are no opportunities for interference as all other pathways are nonresonant with the first interaction. 
A possible mechanism for interfering output in the SR-SIVE-I pathway are cascaded second order processes (SFG, DFG). \cite{RN243, RN300}
In isotropic media, SFG and DFG vanish under the electric dipole approximation,\cite{Zhu87} so cascades are unlikely to inhibit output in centrosymmetric media.
In media where SFG and DFG are allowed, e.g. surfaces or noncentrosymmetric media, second order cascades may become important. 
Generally, for nonlinear spectroscopies which share the same phasematching conditions as SR-SIVE (e.g., DOVE), it is useful to note that SR-SIVE-IR-I is not canceled, whereas SR-SIVE-IR-II and SR-SIVE-Raman permanently cancel.

@@

\section{SIVE of cyanocobalamin}

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{cob.png}
	\caption{
		(a,c) Singly Resonant SIVE (SR-SIVE) and (b,d) Doubly Resonant SIVE (DR-SIVE) response of CNCbl using the $-\vec{k}_1 + 2\vec{k}_2$ and $\vec{k}_1 + 2\vec{k}_2$ phasematching geometries, respectively.
		Here, $\omega_2 = 8000$ cm$^{-1}$.
		$\omega_1$ is stepped in 10 cm$^{-1}$ intervals.
		The spectra are smoothed and power-normalized for frequency dependent power fluctuations during the scan.
		Gray pixels indicate spectral noise.
		The blue trace in (a), (b) corresponds to the spectrum at $\tau_{12}$ = 0 in (c,d), respectively.
		The red trace is an FT-IR spectrum of CNCbl, sourced from NIST. 
		The spectra are normalized to their most intense feature on $\omega \in [1300, 1700]$ cm$^{-1}$.}
	\label{fig:cobsive}
\end{figure*}

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Singly Vibrationally Enhanced (SIVE) spectroscopy, a type of nonlinear hyper-Raman spectroscopy, is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
%alternate name is hyper-(S/D)FG... which is more interesting and descriptive than SIVE imo. lmk your thoughts
We derive selection rules for singly resonant SIVE (SR-SIVE) and demonstrate that SR-SIVE response extracts hyper-Raman polarizabilities ($\beta$).
The SIVE output is shown to be slightly stronger than vibrational sum frequency generation (vSFG), heavily driven by the number density probed by each process. 
The theory of SIVE is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant SIVE (DR-SIVE), akin to resonant hyper-Raman spectroscopy. 
DR-SIVE is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.
\end{abstract}

@@

We investigate properties of doubly resonant SIVE (DR-SIVE) to understand how coupling between vibrational and vibronic states can be measured through DR-SIVE.
DR-SIVE is shown to demand one and two-photon absorptions between the ground and excited state.
DR-SIVE is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}

\section{Results and Discussion}
Singly resonant SIVE (SR-SIVE) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between SR-SIVE spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we make the connections between SIVE and hyper-Raman scattering explicit and remark on the properties of SIVE that become apparent after inspecting the selection rules.

@@

\subsection{Vibrational SR-SIVE Selection Rules}
To make the connection between SR-SIVE and hyper-Raman scattering, we investigate the gross selection rules of SR-SIVE-IR.
By propagating density matrix elements in the steady state limit, the SIVE hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a SIVE experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1934, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
As such, any mode that is SIVE active must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is SIVE active.
This makes SIVE an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
Unlike SFG spectroscopy, which is dependent upon both the infrared and Raman activity of a mode (through symmetry constraints, this demands that the microscopic SFG output from centrosymmetric species vanishes),\cite{Shen90, Cotton} SIVE is a nonlinear, optically upconverted spectroscopy for any infrared active mode.

@@

\subsection{Comparison to Sum Frequency Generation}
Since the hyper-Raman process is effectively a two-photon process, a setup using only two lasers can perform the SIVE experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes.
The other commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique, whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
The SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, this enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole moment.
This follows from simple parity arguments, where inversion centers demand states possess distinct parity selection rules. \cite{RN230}
Macroscopically, inversion symmetry in the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output only depends on surface species.
However, since SIVE spectroscopy is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how SIVE output compares to SFG.

To motivate the application of SIVE spectroscopy, we perform a calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

\subsection{Quantitative SR-SIVE}
With the selection rules of SR-SIVE understood for vibrational spectroscopy, it becomes possible to extract quantitative information from SIVE spectra to compare to incoherent methods and other CMDS methods.
Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
In the method introduced by Levenson and Bloembergen (Bloembergen Interferometry Experiment), an internal standard interferes with the resonant lineshape.
The internal standard is often chosen to be C$_6$H$_6$ or C$_6$D$_6$, due to the strong Raman activity and measured $\chi^{(3)}$ of its ring breathing mode ($\nu_2$ in Herzberg notation). \cite{Levenson1974_2, RN351, RN345}
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Resonant lineshapes are also complicated by amplitude level interference between the sample and its substrate and/or sample cell windows, which must be accounted for to obtain quantitatively correct $\chi^{(3)}$ values. \cite{RN362, RN418}

@@

@@

\section{Conclusions}
Coherent vibrational, hyper-Raman coherent four wave mixing spectroscopies are identified and discussed.
Singly resonant SIVE (SR-SIVE) processes are shown to be the coherent hyper-Raman analogue of infrared active vibrations.
Since SR-SIVE is always non-zero for IR active vibrations, the method could be used as an analytical method to optically upconvert infrared transitions with small transition dipoles.
A method for measuring hyper-Raman polarizabilities, $\beta_{ijk}$, through SIVE spectroscopy was identified and discussed.
Extracting the magnitude and sign of $\beta_{ijk}$ using SIVE methods can help assess the quality of theoretical methods which calculate hyper-Raman spectra. 
Doubly resonant SIVE (DR-SIVE) processes show promise as a probe of vibronic coupling in molecular samples.
Unlike SR-SIVE, DR-SIVE methods can linenarrow inhomogeneously broadened vibronic modes.
